The importance of a complete statistical description of dependence between variables for ecology and related fields

### 2018-06-30 Paper.Rmd
adding biv-soilCN results_Q1

@@

```{r read_BivMS_soilCN_data,echo=F}
BivMS_res_Loecke<-readRDS("./Results/fitting_results/BivMS_soilCN.RDS")
```

@@

The test of indepedence gave $p=$ `r BivMS_res_Loecke$IndepTestRes` (to within the precision available 
from `BiCopIndTest` in the `VineCopula` package). We fitted several copulas
and got AIC values and upper- and lower-tail dependency estimates based on the fitted copulas (Table \ref{SM-tab_soilCNfit}). The table shows that several copulas are much better supported (higher AIC and BIC weights) than the Normal copula, answering the first part of question 1 for these data.

The model-averaged (AIC weights used for model averaging) lower-tail dependence statistic was `r BivMS_res_Loecke$relLTdep_AICw` and the model-averaged upper-tail dependence statistic was `r BivMS_res_Loecke$relUTdep_AICw`. These values are distinct from what a Normal copula would give (i.e., 0), answering the second part of question 1 for these data. Furthermore, the difference of these two quantities (lower minus upper), `r BivMS_res_Loecke$relLTdep_AICw-BivMS_res_Loecke$relUTdep_AICw`, was different from 0, answering the third part of question 1.

@@

As described in \nameref{Methods}, each of our statistics, as computed for the 
ranges $0-0.1$, $0.1-0.2$, ..., $0.9-1$ (with $10$ number of bins), was compared to the distributions 
of values of the same statistic with the same ranges computed on 1000 normal-copula 
surrogate datasets with similar Kendall or Spearman correlations, in different 
runs for Kendall and Spearman (Fig. \ref{SM-fig_soilCN_nonparam}). 
<!--
Because the correlation 
used in the $\cor$ statistic is the Spearman correlation, Spearman-preserving surrogates
should be used, so the top-right panel of Fig. \ref{SM-fig_soilCN_nonparam}
is more reliable than the top left.
--> 
Results indicate that tail dependence of 
data is stronger in both the lower and upper tails than would be expected under a 
null hypothesis of a normal copula, answering the second part of question 1. The dependence between C and N is also 
weaker in the middle of the distributions
than would be expected based on a normal-copula null model.

@@

### BMR data \label{Ex_BMR}

```{r read_BivMS_BMR_data,echo=F}
BivMS_BMR<-readRDS("./Results/BMR_results/BivMS_lnBMR_vs_lnmass.RDS")
```

#### Model selection approach \label{Ex_BMR_Model_selection}

The test of indepedence gave $p=$ `r BivMS_BMR$IndepTestRes` (to within the precision available from `BiCopIndTest` in the `VineCopula` package). We fitted several copulas
and got AIC values and upper- and lower-tail dependency estimates based on the fitted copulas (Table \ref{SM-tab_BMRfit}). The table shows that BB6 and BB8 are much better supported (higher AIC and BIC weights) than the Normal copula, answering the first part of question 1 for these data.

The model-averaged (AIC weights used for model averaging) lower-tail and upper-tail dependence statistic was `r BivMS_BMR$relLTdep_AICw - BivMS_BMR$relUTdep_AICw`, indicating more upper tail dependence. We also tested that lowest-AIC copula BB8 was a good-enough fit, giving $p$-value `r BivMS_BMR$GofRes_CvM` (to within the precision available from `BiCopGofTest` in the `VineCopula` package) in the Cramer-von Mises-based test of the goodness of its fit, and $p$-value `r BivMS_BMR$GofRes_KS` in the Kolmogorov-Smirnov-based test of the goodness of fit. The next section looks at tail dependence in a non-parametric way.

#### Non-parametric analysis of tail dependence \label{Ex_BMR_NPA}

As described in \nameref{Methods}, each of our statistics, as computed for the ranges $0-0.1$, $0.1-0.2$, ..., $0.9-1$ (with $10$ number of bins), was compared to the distributions of values of the same statistic with the same ranges computed on 1000 normal-copula surrogate datasets with similar Kendall or Spearman correlations, in different runs for Kendall and Spearman (Fig. \ref{SM-fig_BMR_nonparam}).

Results indicate that dependence of data is different in some bins than would be expected under a 
null hypothesis of a normal copula, answering the second part of question 1. Table \ref{SM-tab_BMR_asym} indicates stronger upper tail dependence for the BMR data copula which is significantly (95% CI) different from that of the symmetric normal copula and thus consistent with model selection results.

### Anolis data \label{Ex_contrasts}

```{r read_BivMS_contrasts_67_data,echo=F}
BivMS_contrasts_67<-readRDS("./Results/Anolis_results/contrasts_csv_results/BivMS_contrasts_67.RDS")
```

#### Model selection approach \label{Ex_contrasts_Model_selection}

The test of indepedence gave $p=0$ <!--`r BivMS_contrasts_67$IndepTestRes` this is e-15 could not compile latex--> (to within the precision available 
from `BiCopIndTest` in the `VineCopula` package). We fitted several copulas
and got AIC values and upper- and lower-tail dependency estimates based on the fitted copulas (Table \ref{SM-tab_contrasts_67fit}). The table shows that several copulas (mostly J and SC) are much better supported (higher AIC and BIC weights) than the Normal copula, answering the first part of question 1 for these data.

The model-averaged (AIC weights used for model averaging) lower-tail dependence statistic was `r BivMS_contrasts_67$relLTdep_AICw` and the model-averaged upper-tail dependence statistic was `r BivMS_contrasts_67$relUTdep_AICw`. These values are distinct from what a Normal copula would give (i.e., 0), answering the second part of question 1 for these data. Furthermore, the difference of these two quantities (lower minus upper), `r BivMS_contrasts_67$relLTdep_AICw-BivMS_contrasts_67$relUTdep_AICw`, was different from 0, answering the third part of question 1.

These results do convincingly show non-Normal copula structure, giving $p$-value `r BivMS_contrasts_67$GofRes_CvM` (to within the precision available from `BiCopGofTest` in the `VineCopula` package) in the Cramer-von Mises-based test of the goodness of its fit, and $p$-value `r BivMS_contrasts_67$GofRes_KS` in the Kolmogorov-Smirnov-based test of the goodness of fit. The next section looks at tail dependence in a non-parametric way.

#### Non-parametric analysis of tail dependence \label{Ex_contrasts_67_NPA}

As described in \nameref{Methods}, each of our statistics, are computed for $(l_b,u_b)$ in the 
ranges $0-0.25$, $0.25-0.5$, $0.5-0.75$, $0.75-1$ (with $4$ number of bins). Anolis data set has lesser number of data compared to other two bivariate data set, so fewer number of bins were chosen to measure tail-dependence.
Those statistics were compared to the distributions of values of the same statistic with the same ranges computed on 1000 normal-copula surrogate datasets with similar Kendall or Spearman correlations, in different 
runs for Kendall and Spearman (Fig. \ref{SM-fig_contrasts_67_nonparam}).

Results indicate that tail dependence of 
data is stronger in both the lower and upper tails than would be expected under a 
null hypothesis of a normal copula, answering the second part of question 1.

We also tested statistically for asymmetry of tail dependence, which seems visible in Fig. \ref{SM-fig_raw_cop_AVG_humerus_ltoe}(B). Results (Table \ref{SM-tab_contrasts_67_asym}) show that upper-tail dependence was, indeed, significantly stronger than lower-tail dependence in these data, answering the third part of question 1.

@@

As described in Methods, model selection among copulas was done separately for each pair of the 10 
sampling locations, thus results are presented in $10 \times 10$ matrix format with rows and columns 
corresponding to locations. Table \ref{SM-tab_count_bestcop} shows that, for the large majority of location 
pairs, some non-normal copula was a better fit, according to AIC, than was the normal copula, answering the first part of 
question 1 for these data. Comparing the 
AIC weight of the best-fitting (lowest AIC) copula family for each pair of sampling locations (Table \ref{SM-tab_count_AICw}) 
to the AIC weight 
of the normal copula family for the same pair (Table \ref{SM-tab_count_AICw_normal}) confirms that the degree of support for normal copulas was usually not
comparable to the degree of support for other families.
Goodness of fit tests based on a Cramer-von Mises statistic (Table \ref{SM-tab_count_p_CvM}) in every case 
failed to reject the hypothesis that the AIC-best copula family was also an objectively adequate description of 
the data, and likewise for goodness of fit tests based on a Kolmogorov-Smirnov statistic (Table \ref{SM-tab_count_p_KS}),
demonstrating that the collection of copula families we used was sufficiently broad that the copulas of 
all pairs of locations could be characterized by at least one of them. In all cases at least `r min((RES_aphid_count$gfc_numBS_success/RES_aphid_count$gfc_numBS),na.rm=T)*100` percent of bootstrap randomizations used for goodness of fit tests successfully lead to values of the statistic for comparison with its value on the real data.

@@

The difference between the model-averaged lower-tail dependence statistic and the model-averaged upper-tail dependence statistic (lower minus upper), shown in Table \ref{SM-tab_count_LTmUT}, is positive for all but a few locations pairs (`r sum((RES_aphid_count$LTdep_AICw-RES_aphid_count$UTdep_AICw)>0,na.rm=T)` out of `r (dim(RES_aphid_count$gfc_p_CvM)[1]^2-dim(RES_aphid_count$gfc_p_CvM)[1])-2*RES_aphid_count$num_indep_loc_pair`), answering the third part of question 1.

@@

Mean values of our statistics ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, $\Dsq_u$, 
$\cor_l-\cor_u$, $\Ps_l-\Ps_u$, $\Dsq_u-\Dsq_l$), as well as mean Kendall and Spearman correlations, were 
computed across all pairs of sampling locations, and confidence intervals of these means were computed via a spatial
resampling scheme, as described in Methods (Table \ref{SM-tab_count_resamp}). Confidence intervals of
$\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, and $\Dsq_u$ always excluded zero and mean values were positive,
indicating tail dependence. Confidence intervals of $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, $\Dsq_u-\Dsq_l$ also always 
exluded $0$ and were positive, indicating asymmetry of tail dependence, a feature that data with normal copula 
structure do not have, providing further evidence answering the third part of question 1.

Plotting Spearman or Kendall correlation between aphid abundance time series for pairs of locations as a function of great circle distance between the locations (Fig. \ref{SM-fig_aphid_count} A,B) showed the typical decline of correlation with 
distance between sampling locations that has been seen for numerous other datasets in earlier work. Plotting 
$\cor_l$, $\cor_u$, $\Ps_l$, and $\Ps_u$  against distance between pairs of locations shows these quantities also decline with distance (Fig. \ref{SM-fig_aphid_count} C-F). Plotting $\Dsq_u$ and  $\Dsq_l$ against distance between pairs of locations shows
these quantities increase with distance (as expected, given the opposite interpretation these two statistics have compared to $\cor_u$, $\Ps_u$ and $\cor_l$, $\Ps_l$, respectively). Plotting $\cor_l - \cor_u$, $\Ps_l - \Ps_u$, and $\Dsq_u - \Dsq_l$ versus distance (Fig. \ref{SM-fig_aphid_count} I-K) did not show a decline with distance, but did illustrate visually the statistical
result of Table \ref{SM-tab_count_resamp} that averages of these quantities across all pairs of sampling locations
were significantly greater than zero, demonstrating asymmetry of tail dependence. Code for these analyses is in [`NonParamStat.R`](https://github.com/sghosh89/BIVAN/blob/master/NonParamStat.R) in the [`BIVAN`](https://github.com/sghosh89/BIVAN) package.

### First flight date data for the leaf-curling plum aphid, *Brachycaudus helichrysi* \label{Ex_ff}

```{r read_RES_aphid_ff,echo=F}
RES_aphid_ff<-readRDS("./Results/fitting_results/AphidCopulaFit_selecloc_firstflight_species_11.RDS")
```

#### Model selection approach \label{Ex_ff_Model_selection}

As in section \nameref{Ex_count_Model_selection}, results are presented in $10 \times 10$ matrix format with rows
and columns corresponding to locations. Table \ref{SM-tab_ff_bestcop} shows that, for the large majority of location pairs, some non-normal copula was a better fit, according to AIC, than was the normal copula, answering the first part of question 1 for these data. Comparing the AIC weight of the best-fitting (lowest AIC) copula family for each pair of sampling locations (Table \ref{SM-tab_ff_AICw}) to the AIC weight of the normal copula family for the same pair (Table \ref{SM-tab_ff_AICw_normal}) confirms that the degree of support for normal copulas was usually not comparable to the degree of support for other families. Goodness of fit tests based on a Cramer-von Mises statistic (Table \ref{SM-tab_ff_p_CvM}) usually failed to reject the 
hypothesis that the AIC-best copula family was also an objectively adequate description of the data, and likewise for 
goodness of fit tests based on a Kolmogorov-Smirnov statistic (Table \ref{SM-tab_ff_p_KS}). In all cases at least `r min((RES_aphid_ff$gfc_numBS_success/RES_aphid_ff$gfc_numBS),na.rm=T)*100` percent of bootstrap randomizations used for
goodness of fit tests successfully lead to values of the statistic for comparison with its value on the real data.

The model-averaged (averaging performed using AIC weights) lower-tail dependence statistic and 
the model-averaged (again, AIC based) upper-tail dependence statistic are in 
Tables \ref{SM-tab_ff_LTdep_AICw} and \ref{SM-tab_ff_UTdep_AICw}, respectively. These values were exclusively positive, thereby differing from what a normal copula would give (i.e., $0$), and answering the second part of question 1 for these data - these data have greater tail dependence (lower and upper) than would be expected from a normal-copula null hypothesis.

The difference between the model-averaged lower-tail dependence statistic and the 
model-averaged upper-tail dependence statistic (lower minus upper), shown in 
Table \ref{SM-tab_ff_LTmUT}, was negative for `r sum((RES_aphid_ff$LTdep_AICw-RES_aphid_ff$UTdep_AICw)<0,na.rm=T)` 
out of `r (dim(RES_aphid_ff$gfc_p_CvM)[1]^2-dim(RES_aphid_ff$gfc_p_CvM)[1])-2*RES_aphid_ff$num_indep_loc_pair` location pairs, answering the third part of question 1. We note this result is opposite the analogous result for aphid abundance
data for the green spruce aphid, i.e., whereas those abundance data showed stronger lower tail dependendence than upper tail dependence, first flights for the leaf-curling plum aphid showed stronger upper tail dependence than lower tail dependence.

#### Non-parametric analysis of tail dependence  \label{Ex_ff_NPA}

Mean values of our statistics ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, $\Dsq_u$, 
$\cor_l-\cor_u$, $\Ps_l-\Ps_u$, $\Dsq_u-\Dsq_l$), as well as mean Kendall and Spearman 
correlations, were computed across all pairs of sampling locations, and confidence intervals of these means were 
computed via a spatial resampling scheme, as described in \nameref{Methods} (Table \ref{SM-tab_ff_resamp}).
Confidence intervals of $\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, and $\Dsq_u$ always excluded 
zero and mean values were positive, indicating tail dependence. Confidence intervals of $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, $\Dsq_u-\Dsq_l$ also always exluded $0$ and were negative, indicating asymmetry of tail dependence, a feature 
that data with normal copula structure do not have, providing further evidence answering the third part of question 1.
The contrast here with the analogous results for abundances of the green spruce aphid is again illuminating.

Plotting Spearman or Kendall correlations between first flight time series for pairs of locations as a function of great circle distance between the locations (Fig. \ref{SM-fig_aphid_ff} A, B) showed the typical (though slight, in this case) decline of correlation with increasing distance. Plotting $\cor_l$, $\cor_u$, $\Ps_l$, and $\Ps_u$ against distances between pairs of locations shows these quantities also appear in most cases to decline slightly with distance (Fig. \ref{SM-fig_aphid_ff} C-F).
$\Dsq_u$ and $\Dsq_l$ appear to slightly increase with distance, as expected given their
opposite meaning (Fig. \ref{SM-fig_aphid_ff} G-H). 
Plotting $\cor_l - \cor_u$, $\Ps_l - \Ps_u$, and $\Dsq_u - \Dsq_l$ against distance (Fig. \ref{SM-fig_aphid_ff} I-K) did not show a decline with distance, but did demonstrate visually the statistical result of Table \ref{SM-tab_ff_resamp} that averages of these quantities across all pairs of sampling locations were significantly less than zero, showing asymmetry of tail dependence.

### Abundance data for *Ceratium furca* in UK seas \label{Ex_pns}

```{r read_RES_plankton_north_sea,echo=F}
RES_plankton_north_sea<-readRDS("./Results/fitting_results/Plankton_North_Sea_CopulaFit_selecloc_species_16.RDS")
```

#### Model selection approach \label{Ex_pns_Model_selection}

As described in Methods and as in previous sections, model selection among copulas was done separately for 
each pair of sampling locations. There were $14$ sampling locations, so results are presented in $14 \times 14$ matrix format, with rows and columns corresponding to locations. Table \ref{SM-tab_pns_bestcop} shows that, for the large majority of location pairs, some non-normal copula was a better fit, according to AIC, than was the normal copula, answering the first part of question 1 for these data. Comparing the AIC weight of the lowest-AIC copula family for each pair of sampling locations (Table \ref{SM-tab_pns_AICw}) to the AIC weight of the normal copula family for the same pair (Table \ref{SM-tab_pns_AICw_normal}) 
confirms that the degree of support for normal copulas was usually not comparable to the degree of support for other families.
Goodness of fit tests based on a Cramer-von Mises statistic (Table \ref{SM-tab_pns_p_CvM}) usually failed to reject the hypothesis that the AIC-best copula family was also an objectively adequate description of the data, and likewise for goodness of fit tests based on a Kolmogorov-Smirnov statistic (Table \ref{SM-tab_pns_p_KS}). In all cases at least
`r min((RES_plankton_north_sea$gfc_numBS_success/RES_plankton_north_sea$gfc_numBS),na.rm=T)*100` percent of bootstrap randomizations used for goodness of fit tests successfully lead to values of the statistic for comparison with its value on the real data.

The model-averaged (averaging performed using AIC) lower-tail dependence statistic and the model-averaged (again, AIC based) upper-tail dependence statistics are in Tables \ref{SM-tab_pns_LTdep_AICw} and \ref{SM-tab_pns_UTdep_AICw}, respectively. These values were exclusively positive, thereby
differing from what a normal copula would give (i.e., 0), and answering the second part of question 1 for these data - these data
have greater tail dependence (lower and upper) than would be expected from a normal-copula null hypothesis.

The difference between the model-averaged lower-tail dependence statistic and the model-averaged upper-tail dependence statistic
(lower minus upper), shown in Table \ref{SM-tab_pns_LTmUT}, was positive for some
location pairs and negative for others, but was positive for 
`r sum((RES_plankton_north_sea$LTdep_AICw-RES_plankton_north_sea$UTdep_AICw)>0,na.rm=T)` out of `r (dim(RES_plankton_north_sea$gfc_p_CvM)[1]^2-dim(RES_plankton_north_sea$gfc_p_CvM)[1])-2*RES_plankton_north_sea$num_indep_loc_pair`
locations pairs. Thus lower tail dependence tended to be stronger than upper tail dependence, answering the third part of question 1.

#### Non-parametric analysis of tail dependence  \label{Ex_pns_NPA}

Mean values of our statistics, as well as mean Kendall and Spearman
correlations, were computed across all pairs of sampling locations, and confidence intervals of these means were computed via a
spatial resampling scheme, as described in Methods (Table \ref{SM-tab_pns_resamp}).
Confidence intervals of $\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, $\Dsq_u$
always excluded $0$ and mean values were positive, indicating tail dependence.
Confidence intervals of $\cor_l-\cor_u$, $\Ps_l-\Ps_u$ and $\Dsq_u-\Dsq_l$
also all were strictly positive, indicating asymmetry of tail dependence, a feature that data with normal copula
structure do not have, providing further evidence answering the third part of question 1.

Plotting Spearman or Kendall correlations between time series for pairs of locations as a function of great circle distance
between the locations (Fig. \ref{SM-fig_pns} A, B) showed the typical decline of correlation with increasing
distance. Plotting $\cor_l$ , $\cor_u$ , $\Ps_l$, and $\Ps_u$ against distances between pairs of locations shows these quantities also appear to decline with distance (Fig. \ref{SM-fig_pns} C-F). $\Dsq_u$ and $\Dsq_l$ appear to increase with distance, as expected given their opposite meaning (Fig. \ref{SM-fig_pns} G-H). Plotting $\cor_l - \cor_u$, $\Ps_l - \Ps_u$, and $\Dsq_u - \Dsq_l$ against distance (Fig. \ref{SM-fig_pns} I-K) did not show a decline with distance, but did demonstrate visually the 
statistical result of Table \ref{SM-tab_pns_resamp} that averages of these 
quantities across all pairs of sampling locations were significantly greater 
than zero, showing asymmetry of tail dependence.

### Methane flux data \label{Ex_methane}

```{r read_RES_methane,echo=F}
RES_methane<-readRDS("./Results/fitting_results/MethaneFluxFit_selecloc_data_pt_thrs_50.RDS")
```

#### Model selection approach \label{Ex_methane_Model_selection}

As described in Methods and as in previous sections, model selection among copulas was done separately for 
each pair of sampling locations. There were $13$ sampling locations, so results are presented in 
$13 \times 13$ matrix format, with rows and columns corresponding to locations. 
Table \ref{SM-tab_methane_bestcop} shows that, for the large majority of location pairs, some 
non-normal copula was a better fit, according to AIC, than was the normal copula, answering the 
first part of question 1 for these data. Comparing the AIC weight of the lowest-AIC copula 
family for each pair of sampling locations (Table \ref{SM-tab_methane_AICw}) to the AIC weight of
the normal copula family for the same pair (Table \ref{SM-tab_methane_AICw_normal}) 
confirms that the degree of support for normal copulas was usually not comparable to the degree of support for other families.
Goodness of fit tests based on a Cramer-von Mises statistic (Table \ref{SM-tab_methane_p_CvM}) usually failed to reject the hypothesis that the AIC-best copula family was also an objectively adequate description of the data, and likewise for goodness of fit tests based on a Kolmogorov-Smirnov statistic (Table \ref{SM-tab_methane_p_KS}). In all cases at least
`r min((RES_methane$gfc_numBS_success/RES_methane$gfc_numBS),na.rm=T)*100` percent of bootstrap randomizations used for goodness of fit tests successfully lead to values of the statistic for comparison with its value on the real data.

The model-averaged (averaging performed using AIC) lower-tail dependence statistic and the model-averaged (again, AIC based) upper-tail dependence statistics are in Tables \ref{SM-tab_methane_LTdep_AICw} and \ref{SM-tab_methane_UTdep_AICw}, respectively. These values were exclusively positive, thereby
differing from what a normal copula would give (i.e., 0), and answering the second part of question 1 for these data - these data
have greater tail dependence (lower and upper) than would be expected from a normal-copula null hypothesis.

The difference between the model-averaged lower-tail dependence statistic and the model-averaged upper-tail dependence statistic
(lower minus upper), shown in Table \ref{SM-tab_methane_LTmUT}, was positive for some
location pairs and negative for others, but was positive for 
`r sum((RES_methane$LTdep_AICw-RES_methane$UTdep_AICw)>0,na.rm=T)` out of `r (dim(RES_methane$gfc_p_CvM)[1]^2-dim(RES_methane$gfc_p_CvM)[1])-2*RES_methane$num_indep_loc_pair`
locations pairs. Thus lower tail dependence tended to be stronger than upper tail dependence, answering the third part of question 1.

#### Non-parametric analysis of tail dependence  \label{Ex_methane_NPA}

Mean values of our statistics, as well as mean Kendall and Spearman
correlations, were computed across all pairs of sampling locations, and confidence intervals of these means were computed via a
spatial resampling scheme, as described in Methods (Table \ref{SM-tab_methane_resamp}).
Confidence intervals of $\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, $\Dsq_u$
always excluded $0$ and mean values were positive, indicating tail dependence. But the confidence intervals of $\cor_l-\cor_u$, $\Ps_l-\Ps_u$ and $\Dsq_u-\Dsq_l$ were marginal, indicating no significant asymmetry of tail dependence.

Plotting Spearman or Kendall correlations between time series for pairs of locations as a function of great circle distance
between the locations (Fig. \ref{SM-fig_methane} A, B) showed the typical decline of correlation with increasing
distance. Plotting $\cor_l$ , $\cor_u$ , $\Ps_l$, and $\Ps_u$ against distances between pairs of locations shows these quantities also appear to decline with distance (Fig. \ref{SM-fig_methane} C-F). $\Dsq_u$ and $\Dsq_l$ appear to increase with distance, as expected given
their opposite meaning (Fig. \ref{SM-fig_methane} G-H). Plotting $\cor_l - \cor_u$, $\Ps_l - \Ps_u$, and $\Dsq_u - \Dsq_l$ against distance (Fig. \ref{SM-fig_methane} I-K) did not show a decline with distance, but did demonstrate visually the 
statistical result of Table \ref{SM-tab_methane_resamp} that averages of these 
quantities across all pairs of sampling locations were more or less equally dispersed around the zero-line, showing no significant asymmetry of tail dependence.

@@

>>>>

To answer Q1, we choose one representative datasets for each category : with a bivariate (soil C and N data) and
a multivariate (green spruce aphid-abundance data) dataset, we summarized our findings as follows.

@@

```{r read_BivMS_soilCN_data,echo=F}
d<-readRDS("Data/RaCA_soilorganicC_soiltotalN_stocks100cm.RDS")
d<-d[,c("SOCstock100","TSNstock100")] #raw soilCN data
BivMS_res_Loecke<-readRDS("./Results/fitting_results/BivMS_soilCN.RDS")
```

As described in the \ref{SM-SoilCN_data}, raw data plot for soil organic C and soil N stocks to 1m of soil depth from the central
pedon of the `r nrow(d)` sites and their corresponding copula plot were presented in Fig. \ref{fig_biv_raw_cop}(A) and (D),respectively.
We carried out model selection analysis and non-parametric statistics to measure tail-dependence (if any) on this copula plot.

@@

The test of indepedence gave $p=$ `r BivMS_res_Loecke$IndepTestRes` (to within the precision available 
from `BiCopIndTest` in the `VineCopula` package) that showed a significant correlation (Kendall's coefficient = `r round(BivMS_res_Loecke$TauVal,2)`). We then fitted several copulas (all $16$ families as mentioned in \nameref{Model_selection})
and got AIC values and upper- and lower-tail dependency estimates based on the fitted copulas (Table \ref{SM-tab_soilCNfit}). 
Based on AIC weights, `r BivMS_res_Loecke$InfCritRes$copname[which.max(BivMS_res_Loecke$InfCritRes$AICw)]` was the best fit. The table shows that several copulas are much better supported (higher AIC and BIC weights) than the Normal copula, answering the first part of question 1 for these data.

The model-averaged (AIC weights used for model averaging) lower-tail dependence statistic was `r BivMS_res_Loecke$relLTdep_AICw` and the model-averaged upper-tail dependence statistic was `r BivMS_res_Loecke$relUTdep_AICw`. These values are distinct from what a Normal copula would give (i.e., 0), answering the second part of question 1 for these data. Furthermore, the difference of these two quantities (lower minus upper), `r BivMS_res_Loecke$relLTdep_AICw-BivMS_res_Loecke$relUTdep_AICw`, was different from 0, answering the third part of question 1. To be precise, this difference being negative, indicated stronger upper-tail dependence than lower.

@@

As described in \nameref{NPA}, each of our statistics, as computed for the 
ranges $0-0.1$, $0.1-0.2$, ..., $0.9-1$ (with $10$ number of bins), was compared to the distributions 
of values of the same statistic with the same ranges computed on 1000 Normal-copula 
surrogate datasets with similar Kendall or Spearman correlations, in different 
runs for Kendall and Spearman (Fig. \ref{SM-fig_soilCN_nonparam}). Each bin in each subplot of Fig. \ref{SM-fig_soilCN_nonparam}
contains a number which indicates how many surrogates out of 1000, are significantly different ($95\%$ confidence interval) from Normal-copula statistics.

Results indicate that tail dependence of data is stronger in both the lower and upper tails than would be expected under a 
null hypothesis of a normal copula, answering the second part of question 1. The dependence between C and N is also 
weaker in the middle of the distributions
than would be expected based on a normal-copula null model.

@@

We summarized our results columnwise in Table \ref{tab_bivar_summary} for all bivariate datasets plotted in Fig. \ref{fig_biv_raw_cop}.

The first $9$ rows of Table \ref{tab_bivar_summary} summarized model selection results [namely, (1) p-value from `BiCopIndTest`, (2) best copula based on minimum AIC, (3) AIC-weight of that best copula, (4) AIC-weight of Normal copula, (5) p-value from `BiCopGofTest` for Cramer-von Mises method, (6) p-value from `BiCopGofTest` for Kolmogorov-Smirnov method, (7) AIC-weights model averaged lower tail dependence, (8) AIC-weights model averaged upper tail dependence, (9)  AIC-weights model averaged (lower-upper) tail dependence].
Last $3$ rows ($10^{th}$ to $12^{th}$) of Table \ref{tab_bivar_summary} summarized non-parametric statistics results for $P$ statistics calculated on two extreme bins ($\Ps_l$ for first bin, $\Ps_u$ for last bin) for each copula data. 
We computed $\Ps_l$, $\Ps_u$, $\Ps_l - \Ps_u$ for the actual copula data and for Kendall-correlation 
preserving $1000$ Normal surrogates and calculated the rank for actual data as the position if all these $1001$ values were 
arranged in ascending order. So, whenever those ranks are $>975$ or $<25$, that indicate a significantly ($95\%$ CI) different 
tail-dependence measures based on Normal-null hypothesis.

For all $3$ bivariate-datasets indicated by each column name (SoilCN, BMR, Anolis) 
in Table \ref{tab_bivar_summary}, Normal copula were always a poor fit compared to 
highest AIC-weight based asymmetric copulas - which answers the first part of Q1. 
Each dataset presented there, has significant upper tail dependence 
compared to Normal as indicated by negative AIC-weights model averaged 
(lower-upper) tail dependence and very low ranks of actual copula 
data for $\Ps_l - \Ps_u$ statistics, which answers the second and third part of Q1
(For detailed results of each dataset see Tables \ref{SM-tab_soilCN_asym} - \ref{SM-tab_contrasts_67_asym} and 
Figs. \ref{SM-fig_soilCN_nonparam} - \ref{SM-fig_contrasts_67_nonparam}).

@@

As described in \nameref{Model_selection}, for multivariable dataset like green-spruce aphid abundance data, model selection among copulas was done separately for each pair of the 10 sampling locations, thus results are presented in $10 \times 10$ matrix format with rows and columns corresponding to locations. Table \ref{SM-tab_count_bestcop} shows that, for the large majority of location pairs, some non-Normal copula was a better fit, according to maximum AIC-weights, than was the Normal copula, answering the first part of question 1 for these data. Comparing the 
AIC weight of the best-fitting (lowest AIC) copula family for each pair of sampling locations (Table \ref{SM-tab_count_AICw}) to the AIC weight of the normal copula family for the same pair (Table \ref{SM-tab_count_AICw_normal}) confirms that the degree of support for Normal copulas was usually not
comparable to the degree of support for other families.
Goodness of fit tests based on a Cramer-von Mises statistic (Table \ref{SM-tab_count_p_CvM}) in every case 
failed to reject the hypothesis that the AIC-best copula family was also an objectively adequate description of 
the data, and likewise for goodness of fit tests based on a Kolmogorov-Smirnov statistic (Table \ref{SM-tab_count_p_KS}), demonstrating that the collection of copula families we used was sufficiently broad that the copulas of all pairs of locations could be characterized by at least one of them. In all cases at least `r min((RES_aphid_count$gfc_numBS_success/RES_aphid_count$gfc_numBS),na.rm=T)*100` percent of bootstrap randomizations used for goodness of fit tests successfully lead to values of the statistic for comparison with its value on the real data.

@@

The difference between the model-averaged lower-tail dependence statistic and the model-averaged upper-tail dependence statistic (lower minus upper), shown in Table \ref{SM-tab_count_LTmUT}, is positive for all but a few locations pairs (`r sum((RES_aphid_count$LTdep_AICw-RES_aphid_count$UTdep_AICw)>0,na.rm=T)` out of `r (dim(RES_aphid_count$gfc_p_CvM)[1]^2-dim(RES_aphid_count$gfc_p_CvM)[1])-2*RES_aphid_count$num_indep_loc_pair`), answering the third part of question 1. Fig. \ref{fig_aphid_raw_cop}(A) and (C) represnt exemplary plots for aphid abundance (green spruce aphid) raw data and its copula form at $(4,11)$ location pair which shows stronger lower tail dependence.

@@

Mean values of our statistics ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, $\Dsq_u$, 
$\cor_l-\cor_u$, $\Ps_l-\Ps_u$, $\Dsq_u-\Dsq_l$), as well as mean Kendall and Spearman correlations, were 
computed across all pairs of sampling locations, and confidence intervals of these means were computed via a spatial
resampling scheme, as described in \nameref{NPA} (Table \ref{SM-tab_count_resamp}). Confidence intervals of
$\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, and $\Dsq_u$ always excluded zero and mean values were positive,
indicating tail dependence. Confidence intervals of $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, $\Dsq_u-\Dsq_l$ also always 
exluded $0$ and were positive, indicating asymmetry of tail dependence (stronger lower tail dependence than upper), a feature that data with Normal copula structure do not have, providing further evidence answering the third part of question 1.

Plotting Spearman or Kendall correlation between aphid abundance time series for pairs of locations as a function of great circle distance between the locations (Fig. \ref{SM-fig_aphid_count} A,B) showed the typical decline of correlation with distance between sampling locations that has been seen for numerous other datasets in earlier work. Plotting $\cor_l$, $\cor_u$, $\Ps_l$, and $\Ps_u$  against distance between pairs of locations shows these quantities also decline with distance (Fig. \ref{SM-fig_aphid_count} C-F). Plotting $\Dsq_u$ and  $\Dsq_l$ against distance between pairs of locations shows these quantities increase with distance (as expected, given the opposite interpretation these two statistics have compared to $\cor_u$, $\Ps_u$ and $\cor_l$, $\Ps_l$, respectively). Plotting $\cor_l - \cor_u$, $\Ps_l - \Ps_u$, and $\Dsq_u - \Dsq_l$ versus distance (Fig. \ref{SM-fig_aphid_count} I-K) did not show a decline with distance, but did illustrate visually the statistical result of Table \ref{SM-tab_count_resamp} that averages of these quantities across all pairs of sampling locations were significantly greater than zero, demonstrating asymmetry of tail dependence, as also seen from model selection approach. Code for these analyses is in [`NonParamStat.R`](https://github.com/sghosh89/BIVAN/blob/master/NonParamStat.R) in the [`BIVAN`](https://github.com/sghosh89/BIVAN) repository.

@@

We summarized our results columnwise in Table \ref{tab_multivar_summary} for all multivariable datasets 
(i.e., abundance data for green spruce aphid from `r nrow(RES_aphid_count$gfc_numBS)` sampling locations, 
firstflight date data for leaf-curling plum aphid from `r nrow(RES_aphid_count$gfc_numBS)` sampling locations, 
abundance data for plankton from UK seas from $14$ sampling locations, 
methane flux data from $13$ sampling locations). We counted for how many location pair (excluding self-comparison)
copula are non-independent.
